DATA 621 - Business Analytics and Data Mining

### 2020-11-02 HW4/HW4_Draft.Rmd
Correct train/test mixup. Work on Sev model

---
title: "DATA 621 - Business Analytics and Data Mining"
subtitle: 'Fall 2020 - Group 2 - Homework #4'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "11/01/2020"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 3
urlcolor: purple
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```

```{r loadData, include=FALSE}
# Load necessary libraries
library(corrplot)
library(ggplot2)
library(scales)
library(knitr)
library(caret)
library(data.table)
# Set master seed
set.seed(71554)

@@

setattr(DT$URBANICITY, 'levels', c('Urban', 'Rural'))
```

@@

```{r model1addVars}
# Add OWN variable for model set 1
tstX[, OWN := factor(ifelse(HOME_VAL > 0, 1, 0),
                     levels = c(0, 1), labels = c('No', 'Yes'))]
```

### Dummy Variables
As there are many factor variables, dummy variables will be created. The target
variables will be removed and held to the side so as not to contaminate the
fitting

```{r model1createDummy}
trnFreq <- tstX$TARGET_FLAG
trnSev <- tstX$TARGET_AMT
tstDumFreq <- TARGET_FLAG ~ . - c(INDEX, TARGET_AMT)
```

# BUILD MODELS
## Model Set 1
### Frequency Model
The frequency model will be a binary logistic on the prepared data. It will
start with the complete model and allow for some interactions. It will use both
forward and backward steps to identify which variables to include, using AIC as
the test metric. Using this approach does not allow for cross-validation. The
initial interactions considered will be:

 * `OWN` with `HOME_VAL` as discussed above
 * `HOME_VAL` with `INCOME` as using both may overstate the effect of wealth
 * `KIDSDRIV` with `HOMEKIDS` as there is

# CODE APPENDIX
```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r loadData}
```
```{r missingVal}
```
```{r zeroVal}
```
```{r statsN}
```
```{r graphsN}
```
```{r graphsF}
```
```{r dataClean}
```
```{r trainTestSplit}
```
```{r model1addVars}
```

>>>>

---
title: "DATA 621 - Business Analytics and Data Mining"
subtitle: 'Fall 2020 - Group 2 - Homework #4'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "11/01/2020"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 4
urlcolor: purple
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```

```{r loadData, include=FALSE}
# Load necessary libraries
library(corrplot)
library(ggplot2)
library(scales)
library(knitr)
library(caret)
library(RANN)
library(data.table)
# Set master seed
set.seed(71554)

@@

setattr(DT$URBANICITY, 'levels', c('Urban', 'Rural'))

# Relevel some of the variables to have the default make more sense
# Set default marital status to unmarried
set(DT, NULL, 'MSTATUS', relevel(DT$MSTATUS, 'No'))
# Set default car usage to private
set(DT, NULL, 'CAR_USE', relevel(DT$CAR_USE, 'Private'))
```

@@

```{r model1addVars}
# Add OWN variable for model set 1
trnX[, OWN := factor(ifelse(HOME_VAL > 0, 1, 0),
                     levels = c(0, 1), labels = c('No', 'Yes'))]
tstX[, OWN := factor(ifelse(HOME_VAL > 0, 1, 0),
                     levels = c(0, 1), labels = c('No', 'Yes'))]
```

# BUILD MODELS
## Model Set 1
### Dummy Variables
As there are many factor variables, dummy variables will be created. The target
variables will be removed and held to the side so as not to contaminate the
fitting.

### Frequency Model
The frequency model will be a binary logistic on the prepared data. It will
start with the complete model and allow for some interactions. It will use both
forward and backward steps to identify which variables to include, using AIC as
the test metric. Using this approach does not allow for cross-validation. The
initial interactions considered will be:

 * `OWN` with `HOME_VAL` as discussed in the missing data section
 * `KIDSDRIV` with `HOMEKIDS`
 * A three-way interaction between `HOME_VAL`, `INCOME`, and `BLUEBOOK`
 * `CLM_FREQ` with `OLDCLAIM`
 * `MSTATUS` with `SEX`
   * One sex may be more affected by marriage than the other

```{r model1DummyFreq}
trnFreq <- factor(trnX$TARGET_FLAG, levels = c(0, 1),
                  labels = c("NoClaim", "Claim"))
trnDumFreq <- dummyVars(TARGET_FLAG ~ .  + OWN * HOME_VAL +
                          KIDSDRIV * HOMEKIDS + HOME_VAL * INCOME * BLUEBOOK +
                          CLM_FREQ * OLDCLAIM + MSTATUS * SEX - INDEX -
                          TARGET_AMT,
                        fullRank = TRUE, data = trnX)
trnXFD <- predict(trnDumFreq, newdata = trnX)
```

#### Training
At this point the code will be set up to preprocess the data and train the
model. The two pre-processing steps which will be done will be to check for
near-zero variance between the predictors and then use kNN imputations for
missing values. The actual code may be found in the CODE APPENDIX.

```{r model1RunFreq}
trC <- trainControl(method = 'none',            # Using StepAIC
                    classProbs = TRUE,          # Classification
                    summaryFunction = prSummary # Precision/Recall)
)
set.seed(487)
m1FreqFit <- train(x = trnXFD, y = trnFreq, preProcess = c('nzv', 'knnImpute'),
                   trControl = trC, method = 'glmStepAIC',
                   family = binomial(link = 'logit'), direction = 'both',
                   trace = 1)
kable(summary(m1FreqFit)$coefficients, digits = 3L,
      caption = "Model 1 Frequency Regression Output")
```

#### Coefficient Discussion
Positive coefficients represent an **increased** risk of a claim over the
baseline and negative coefficients represent a **decreased** risk over the
baseline. Most of the coefficients are of the expected direction. For example,
having one's license revoked or living in an urban area is a *positive* risk
factor for claims while higher levels of wealth or education are *negative* risk
factors.

Some interesting observations. `SEX` is not a significant predictor, even at the
10% level, but removing it does not increase the AIC, so it remains. The
hypothesized interaction between `MSTATUS` and `SEX` does lower the deviance,
but not enough to justify the additional parameter.

The predictors with the greatest effect on risk are `URBANICITY` and `INCOME`.
Having the three-way interaction between `HOME_VAL`, `INCOME`, and `BLUEBOOK`,
together with `INCOME` and `BLUEBOOK` alone, obviated the need for stand-alone
`HOME_VAL` and the other two possible interactions.

In terms of variable importance, it can be estimated from the the last trace of
the stepping procedure shown below. The further down the table the variable is,
the more dramatic its inclusion or exclusion will be. The top 10 variables in
order would be: `URBANICITY.Rural`, `REVOKED.Yes`, `CAR_USE.Commercial`,
`TRAVTIME`, `CAR_TYPE.Sports Car`, `CLM_FREQ`, `MVR_PTS`, `TIF`, `CAR_TYPE.SUV`,
and `JOB.Blue Collar`.

```
                             Df Deviance    AIC
<none>                            5164.6 5224.6
+ HOMEKIDS                    1   5162.7 5224.7
+ AGE                         1   5163.1 5225.1
- SEX.F                       1   5167.4 5225.4
+ EDUCATION.Masters           1   5163.4 5225.4
+ JOB.Manager                 1   5163.7 5225.7
+ OLDCLAIM                    1   5164.0 5226.0
+ EDUCATION.PhD               1   5164.0 5226.0
+ `HOME_VAL:BLUEBOOK`         1   5164.2 5226.2
+ CAR_AGE                     1   5164.2 5226.2
+ `MSTATUSYes:SEXF`           1   5164.5 5226.5
+ OWN.Yes                     1   5164.5 5226.5
+ YOJ                         1   5164.6 5226.6
+ RED_CAR.yes                 1   5164.6 5226.6
+ `EDUCATION.High School`     1   5164.6 5226.6
+ `INCOME:HOME_VAL:BLUEBOOK`  1   5164.6 5226.6
- `INCOME:BLUEBOOK`           1   5168.8 5226.8
- `INCOME:HOME_VAL`           1   5170.0 5228.0
- JOB.Lawyer                  1   5170.6 5228.6
- JOB.Student                 1   5171.4 5229.4
- `CAR_TYPE.Panel Truck`      1   5173.5 5231.5
- MSTATUS.Yes                 1   5174.8 5232.8
- CAR_TYPE.Van                1   5179.5 5237.5
- BLUEBOOK                    1   5179.9 5237.9
- `OLDCLAIM:CLM_FREQ`         1   5180.0 5238.0
- `JOB.Home Maker`            1   5180.2 5238.2
- HOME_VAL                    1   5180.7 5238.7
- EDUCATION.Bachelors         1   5181.4 5239.4
- INCOME                      1   5182.3 5240.3
- PARENT1.Yes                 1   5182.6 5240.6
- CAR_TYPE.Pickup             1   5183.2 5241.2
- JOB.Professional            1   5186.1 5244.1
- JOB.Clerical                1   5195.4 5253.4
- KIDSDRIV                    1   5197.8 5255.8
- `JOB.Blue Collar`           1   5201.2 5259.2
- CAR_TYPE.SUV                1   5202.5 5260.5
- TIF                         1   5204.4 5262.4
- MVR_PTS                     1   5204.8 5262.8
- CLM_FREQ                    1   5205.4 5263.4
- `CAR_TYPE.Sports Car`       1   5208.4 5266.4
- TRAVTIME                    1   5211.7 5269.7
- CAR_USE.Commercial          1   5230.2 5288.2
- REVOKED.Yes                 1   5233.0 5291.0
- URBANICITY.Rural            1   5612.1 5670.1
```

### Severity Model
In the insurance world, severity is not necessarily fit via linear models, but
is more often fit through comparison of the maximum likelihood fits of various
distributional families to the size of the claims, but the task here is to fit a
linear regression model to estimate future claim size.

#### Training
The observations in the training set which have claim amounts will be extracted
and used to fit the linear regression. Similar to frequency, the model will be
selected via a forward and backward stepwise AIC algorithm starting with every
relevant predictive variable and some selected interactions.

```{r model1DummySev}
trnSev <- trnX$TARGET_AMT[trnX$TARGET_AMT > 0]
trnDumSev <- dummyVars(TARGET_AMT ~ .  + OWN * HOME_VAL +
                          KIDSDRIV * HOMEKIDS + HOME_VAL * INCOME * BLUEBOOK +
                          CLM_FREQ * OLDCLAIM + MSTATUS * SEX - INDEX -
                          TARGET_FLAG,
                        fullRank = TRUE, data = trnX[TARGET_AMT > 0])
trnXSD <- predict(trnDumFreq, newdata = trnX[TARGET_AMT > 0])
```

```{r runModel1Sev}
trC <- trainControl(method = 'none')
set.seed(878)
m1SevFit <- train(x = trnXSD, y = trnSev, preProcess = c('nzv', 'knnImpute'),
                   trControl = trC, method = 'glmStepAIC',
                   family = gaussian, direction = 'both',
                   trace = 0)
kable(summary(m1SevFit)$coefficients, digits = 3L,
      caption = "Model 1 Severity Regression Output")
```

#### Coefficient Disucssion
Hi!

```
                             Df   Deviance   AIC
<none>                          9.5466e+10 31329
- REVOKED.Yes                 1 9.5614e+10 31329
+ JOB.Manager                 1 9.5368e+10 31330
- MSTATUS.Yes                 1 9.5625e+10 31330
+ CAR_AGE                     1 9.5379e+10 31330
- `OLDCLAIM:CLM_FREQ`         1 9.5645e+10 31330
+ `MSTATUSYes:SEXF`           1 9.5392e+10 31330
- CAR_USE.Commercial          1 9.5650e+10 31330
+ SEX.F                       1 9.5399e+10 31330
+ JOB.Professional            1 9.5400e+10 31330
+ CAR_TYPE.Pickup             1 9.5410e+10 31330
+ CAR_TYPE.Van                1 9.5412e+10 31330
+ MVR_PTS                     1 9.5413e+10 31330
+ HOMEKIDS                    1 9.5415e+10 31330
+ CLM_FREQ                    1 9.5431e+10 31331
- `INCOME:HOME_VAL`           1 9.5688e+10 31331
+ RED_CAR.yes                 1 9.5440e+10 31331
- HOME_VAL                    1 9.5694e+10 31331
+ JOB.Student                 1 9.5444e+10 31331
+ EDUCATION.Masters           1 9.5445e+10 31331
+ PARENT1.Yes                 1 9.5452e+10 31331
- OLDCLAIM                    1 9.5709e+10 31331
+ JOB.Clerical                1 9.5455e+10 31331
+ `CAR_TYPE.Panel Truck`      1 9.5455e+10 31331
+ `EDUCATION.High School`     1 9.5456e+10 31331
+ OWN.Yes                     1 9.5457e+10 31331
+ TRAVTIME                    1 9.5458e+10 31331
+ `INCOME:HOME_VAL:BLUEBOOK`  1 9.5460e+10 31331
+ `JOB.Home Maker`            1 9.5460e+10 31331
+ URBANICITY.Rural            1 9.5463e+10 31331
+ `INCOME:BLUEBOOK`           1 9.5464e+10 31331
+ AGE                         1 9.5464e+10 31331
+ `HOME_VAL:BLUEBOOK`         1 9.5465e+10 31331
+ `CAR_TYPE.Sports Car`       1 9.5465e+10 31331
+ YOJ                         1 9.5465e+10 31331
+ `JOB.Blue Collar`           1 9.5466e+10 31331
+ JOB.Lawyer                  1 9.5466e+10 31331
+ KIDSDRIV                    1 9.5466e+10 31331
+ EDUCATION.Bachelors         1 9.5466e+10 31331
+ CAR_TYPE.SUV                1 9.5466e+10 31331
+ `KIDSDRIV:HOMEKIDS`         1 9.5466e+10 31331
+ TIF                         1 9.5466e+10 31331
+ INCOME                      1 9.5466e+10 31331
+ EDUCATION.PhD               1 9.5466e+10 31331
- BLUEBOOK                    1 9.6124e+10 31337
```
# SELECT MODELS
## Model Selection Criteria
For that *classification* algorithm, we will look at accuracy, AUC, and F1, and
select the model which performs best two out of three in those metrics. For the
*regression* algorithm we will look at RMSE, AIC. and MAE (since the values must
be positive) and select the model which performs best two out of three in those
metrics.

### Frequency Models

```{r model1Test}

```

# PREDICTIONS

# CODE APPENDIX
```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r loadData}
```
```{r missingVal}
```
```{r zeroVal}
```
```{r statsN}
```
```{r graphsN}
```
```{r graphsF}
```
```{r corrgram}
```
```{r dataClean}
```
```{r trainTestSplit}
```
```{r model1addVars}
```
```{r model1DummyFreq}
```
```{r model1RunFreq}
```
```{r model1DummySev}
```
```{r model1RunSev}
```
